AppPAL for Android

### 2015-09-23 paper.tex
updates

@@

\begin{abstract}
  It can be difficult to find appropriate mobile apps.
  At work, employers rely on employees enforcing company policies about apps at the workplace correctly.
  Users must judge apps by the information shown to them by the store.
  Whilst some users take time to pick with care others do not.
  Users feel frustrated when they realise what data the apps have access to.
  They want greater control over what data they give away but they do not want to spend time reading permissions lists.
  We present AppPAL: a policy language to enforce policies about apps.
  AppPAL policies use statements from third parties and delegation relationships to give us a rigorous and flexible framework for enforcing and comparing app policies; and the trust relationships surrounding them.
  Finally we show that AppPAL can be used to model user's privacy policies but that most users do not seem to enforce them when picking apps for their phone: a demonstration of the privacy paradox.
\end{abstract}

@@

Finding the right apps can be tricky.
Users need to discover which apps are well written, which are not going to abuse their data
  and to find the apps which suit how they want to use their device.
This can be difficult as it isn't obvious how apps use the data each has access to.

App stores give some information about their apps; descriptions of the app and screenshots as well as review scores.
Android apps show a list of permissions when they're first installed.
Soon new apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use and which permissions to grant must be made by the user.

Not all apps are suitable.
Many \ac{pup} are being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work (a scheme called \ac{byod}).
An employer may wish to restrict which apps their employees can use.
The IT department may set a policy---a series of rules describing what kinds of apps may be used and how---to prevent information leaks.
Some users worry apps will misuse their personal data;
  such a user avoids apps which can access their location, or address book.
They may apply their own personal security policies when downloading and running apps.

These policies can only be enforced by the users continuously making the correct decision when prompted about apps.
This is error-prone.
We believe this can be improved.
An alternative is be to write the policy down and make the computer enforce it.
To implement this we use a logic of authorization---a language designed to express rules about permissible actions.
The policy is written in the logic and enforced by checking the policy is satisfied.

We present AppPAL, an authorization logic for reasoning about apps.
The language is an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool}. and predicates that allow us to decide which apps to run or install.
The language allows us to reason about apps using statements from third parties.
The implementation allows us to enforce the policies on a device.
We can express trust relationships amongst these parties; use constraints to do additional checks.
This lets us enforce deeper and more complex policies than existing tools such as Kirin~\cite{Enck:2009ko}.

Using AppPAL we write policies for work and home, and decide which policy to use using a user's location, or the time of day:
\begin{lstlisting}
"alice" says App isRunnable
  if "home-policy" isMetBy(App)
  where At("work") = false.

"alice" says App isRunnable
  if "work-policy" isMetBy(App)
  where BeforeHourOfDay("17") = true.
\end{lstlisting}
We can delegate policy specification to third parties or roles, and assign principals (entities which speak) to roles:
\begin{lstlisting}
"alice" says "it-department" can-say 0 "work-policy" isMetBy(App).
"alice" says "alice" can-act-as "it-department".
\end{lstlisting}
We can write policies specifying which permissions an app must or must not have by its app store categorization:
    \begin{lstlisting}
"alice" says App isRunnable
   if "permissions-policy" isMetBy(App).
"alice" says "permissions-policy" isMetBy(App)
   if App isAnApp
   where
      category(App, "Photography"),
      hasPermission(App, "LOCATION") = false,
      hasPermission(App, "CAMERA") = true.
    \end{lstlisting}

% Using AppPAL we can say an app is installable:
% \begin{lstlisting}
% "Alice" says "com.rovio.angrybirds" isInstallable.
% \end{lstlisting}
% We can delegate decisions to others:
% \begin{lstlisting}
% "Alice" says "ITDepartment" can-say 0 App isInstallable.
% \end{lstlisting}
% We can run analysis tools, and solve constraints:
% \begin{lstlisting}
% "Claire" says App isMalware
%   if App isAnApp
%   where
%     virusScanner(App) = true.
% \end{lstlisting}
% Constraints allow up to express facts that are true at some times but not others:
% \begin{lstlisting}
% "Emma" says "com.facebook.katana" isInstallable
%   where timeOfDay() > 1700.
% \end{lstlisting}

\noindent
In this paper we have:
\begin{itemize}
  \item
    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
    and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
  \item
    Implemented the AppPAL language as a library on Android and the JVM.
    We show how the language can describe app properties and Android security policies (\autoref{ssec:language}) as well as how to check policies (\autoref{ssec:eval}).
  \item
    Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
\end{itemize}

@@

An employee \emph{Alice} works for her \emph{em\/}ployer \emph{Emma}.
Emma allows Alice to use her personal phone as a work phone, but she has some specific concerns.
\begin{itemize}
  \item Alice shouldn't run any apps that can track her movements.
    The testing labs are at a secret location and it mustn't be leaked.
  \item Apps should come from a reputable source, such as the Google Play Store.
  \item Emma uses an \ac{av} program by McAfee.
    It should check all apps before they're installed.
\end{itemize}

To ensure this policy is met Alice promises to follow it.
She might even sign a document promising never to break the rules within the policy.
This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
Emma's policy could be implemented using existing tools.
\emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}).
AppPAL is designed to build on these tools, but make the trust and delegation relationships explicit and provide a rigorous means for deciding whether an app meets a policy.
Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
These could be used to used to ensure no location data is ever obtained.
Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are never leaked to the web.
Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.

@@

Our research looks at the problem of picking software.
Given there are some apps you want to install and run and others you do not:
  how can you express your preferences in such a way that they can be enforced automatically?
How can we translate policy documents from natural language into a machine checkable form?
How can we show the trust relationships used to make these decisions clearly and precisely?

@@

We propose a change to the ecosystem, shown in \autoref{fig:ecosystem}.
People have policies which are enforced by AppPAL on their devices.
They can be composed with policies from employers or others to create enhanced devices that ensure apps meet the policies of their owners.
The device can make use of vetting services which run tools to infer complex properties about apps.
Users can buy from enhanced stores which ensure the only apps they sell are the apps which meet the explicitly specified store policies.
Developers could decide which stores to sell their apps in on the basis of policies about stores.

@@

"emma" says "google-play" can-say 0
  App isReputable. @\label{lst:corporate_4}@
    \end{lstlisting}
  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       firstnumber=15,
                       basicstyle=\scriptsize\ttfamily,
                       stringstyle=\scriptsize\sffamily,
                       keywordstyle=\scriptsize\slshape]
"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = false.

@@

In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isRunnable} (a predicate).
She allows her to delegate the decision further if she chooses (\code{can-say inf}).
Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
If Emma can be convinced all these policies are met then she will say the \code{App isRunnable}.
In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isReputable};
  with \code{"google-play"} specified as the decider of what is buyable or not.
This time Google is not allowed to delegate the decision further (\code{can-say 0}).
In other words Google is not allowed to specify Amazon as a supplier of apps as well.
Google must say what is buyable directly for Emma to accept it.
Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
Here we use a constraint.
When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
Only if this returns \code{false} will the policy be met.
To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.

@@

AppPAL is implemented as a library for Android and Java.
The parser is implemented using ANTLR4.
The structure of an AppPAL assertion can be seen in \autoref{fig:assertion}.

\begin{figure}
  \newcommand{\bracetext}[1]{\text{\sffamily #1}}
  \newcommand{\smalltext}[1]{\text{\ttfamily\scriptsize #1}}
  \centering
  \begin{minipage}{0.49\linewidth}
    \begin{equation*}\small
      \begin{array}{r l}\footnotesize
        \overbrace{\smalltext{`user'}}^{\bracetext{speaker}} &
        \smalltext{ says }\overbrace{\overbrace{\smalltext{ App }}^{\bracetext{subject}}\overbrace{\smalltext{ isRunnable}}^{\bracetext{predicate}}}^{\bracetext{fact}} \\
        & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
        & \overbrace{\smalltext{ where hasPermission(App, `INTERNET') = true}}^{\bracetext{constraint}}.
      \end{array}
    \end{equation*}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}
  \newcommand{\nonterminal}[1]{$\langle$#1$\rangle$}
  \newcommand{\terminal}[1]{\textbf{#1}}
  \begin{tabular}{r c l}
    \footnotesize
    \nonterminal{Assertion} & $\coloneqq$ & \nonterminal{E} \terminal{says} \nonterminal{Fact} \\
                            &             & \hspace{1em}(\terminal{if} (\nonterminal{Fact}\terminal{,})+)?\terminal{.} \\
                            &             & \hspace{1em}(\terminal{where} \nonterminal{Constraint})? \\
    \nonterminal{Fact}      & $\coloneqq$ & \nonterminal{E} (\terminal{isRunable} $\vert$ $\ldots$) \\
                            & $\vert$     & \nonterminal{E} \terminal{can-say} (\terminal{0} $\vert$ \terminal{inf}) \nonterminal{Fact} \\
                            & $\vert$     & \nonterminal{E} \terminal{can-act-as} \nonterminal{E} \\
    \nonterminal{E}         & $\coloneqq$ & \terminal{Variable} $\vert$ \terminal{`constant'}
  \end{tabular}
  \end{minipage}
  \caption{Structure and simplified grammar of an AppPAL assertion.}
  \label{fig:assertion}
\end{figure}

In Becker~\etal's work~\cite{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
With AppPAL we make explicit our predicates and how they relate to Android.
AppPAL policies can make use of the predicates and constraints in \autoref{tab:predicates}.
Additional predicates can be created in the policy files, but adding or modifying constraints required a code change\footnote{The Android version of the \texttt{hasPermission} constraint, for example, uses the Android package manager to determine what permissions an app requests, but the Java version uses the Android platform tools.  This requires a small code change.}

\begin{table}
\begin{tabulary}{\linewidth}{l L}
  \toprule
  Name & Description \\
  \midrule
  App \texttt{isRunnable} & Says an app can be run. \\
  App \texttt{isInstallable} & Says an app can be installed. \\
  App \texttt{isAnApp} & Tells AppPAL that an app exists. \\
  Policy \texttt{isMetBy(}App\texttt{)} & Says a specific policy is met by an app.  This is used to split policies into smaller components which can be reused and composed. \\
  \texttt{hasPermission(}App, Permission\texttt{)} & Constraint for testing whether an app has been granted a permission. \\
  \texttt{BeforeHourOfDay(}time\texttt{)} & Constraint used to test whether we're before an hour in the day. \\
  Tool\texttt{Check(}App, Property) &  General form of a constraint used to run a static analysis tool over an app. \\
  \bottomrule
\end{tabulary}
\caption{Typical AppPAL predicates and constraints}
\label{tab:predicates}
\end{table}

% \begin{description}
%   \item[\texttt{App isRunnable}]
%     Used to indicate that an App meets the install policy for the device.
%     Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.
%
%   \item[\texttt{App isAnApp}]
%     Expresses a simple typing relations.
%     AppPAL, and SecPAL, have a safety condition that all variable mentioned in the fact, are also mentioned in the condition body (as is the case for Datalog and other logic languages).
%     These typing relations allow an appeal to ground facts to be made so the safety condition is satisfied.
%     This typically is necessary when an assertion has a constraint but no body conditions: we add the trivial typing statement to the body to ensure the assertion is accepted by AppPAL.
%
%   \item[\texttt{Policy isMetBy(App)}]
%     Policies let you split app behaviour into sub-policies.
%     For example in \autoref{sec:idea} we showed how Emma's installation policy could be written using three \texttt{isMetBy} statements.
%     Splitting policies allows greater control of how each one is checked.
%     We can delegate checking a policy to an expert using the \code{can-say} statement.
%     We can specify more detailed checks using the \code{can-act-as} statements.
%
%   \item[\texttt{Policy shownIsMetBy(Evidence, App)}]
%     A variant of the \texttt{isMetBy} statement that allows some evidence (a proof) to be given that shows the policy is met.
% \end{description}

@@

In \autoref{sec:problem} and \autoref{sec:idea} we described a \emph{no-tracking-policy} to prevent a user's location being leaked.
In Emma's policy we checked this using the app's permissions.
If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
Some apps may want to access this data, but may not leak it.
We could use a taint analysis tool to detect this (e.g.~Taintdroid~\cite{Fritz:2013vi}).
Our policy now becomes:

@@

"emma" says "location-taint-analysis" isMetBy(App)
  if App isAnApp
  where
    taintDroidCheck(App, "Location", "Internet") = false.
\end{lstlisting}

@@

AppPAL has the same policy inference rules as SecPAL.
We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based checking algorithm, as no DatalogC library exists for Android.
We have implemented the rules directly in Java.
Pseudo-code is given in \autoref{fig:pseudocode}.

Like Becker~\etal~we make use of an assertion context to store known statements.
We also cache intermediate results, and ground facts to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
This gives us greater control of the evaluation and how the assertion context is created.
For example, when checking the \code{isAnApp} predicate;
  we can fetch the assertion that the subject is an app based on the app in question.
% Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
%   it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.

% In SecPAL there is no type.
% All entities are either constants (implemented as strings) or variables.
% There is no differentiation made between a \emph{voiced} constant (who says assertions),
% and a \emph{subject} (the subject of a verb-phrase, or predicate argument).
% In AppPAL the subjects are usually apps or policies.
% Apps do not utter facts;
%   so there is no need to check them as potential speakers of \emph{can-say} statements.
% Similarly if a constant never appears as a subject of a statement, as most speakers do not,
%   there is no need to check it as a possible subject for a \emph{can-act-as} statement.
% We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
% This means we need to search less constants when attempting to replace a variable, speeding evaluation.

@@

% \subsection{Policy Examples}
% \label{ssec:idioms}
%
% When writing policies some schemes often come up.
% We give examples of policies and queries and show how they can be implemented in AppPAL.
%
% \begin{description}
%   \item[Policies for home and work]
%     In \autoref{sec:introduction} we said Alice might like to have separate rules for home and work.
%     Emma may insist that her policies come into effect whenever Alice is at work, but is okay with Alice breaking them after 5~pm or when she isn't at work.
%     To implement this we add rules that use constraints to decide whether an app should be run.
%     Alice already has the Emma's work policy as the default.
%     Now she just needs to add an exception to the rule.
%     \begin{lstlisting}
% "alice" says App isRunnable
%   if "home-policy" isMetBy(App)
%   where TimeOfDay() > 17.00.
%
% "alice" says App isRunnable
%   if "home-policy" isMetBy(App)
%   where At("work") = false.
%     \end{lstlisting}
%
%   \item[App white-listing by an IT department]
%     Inside Emma's company employees in the IT department may white-list certain apps as runnable.
%     Emma delegates to them to decide if an app is runnable.
%     \begin{lstlisting}
% "emma" says "it-department" can-say 0 App isRunnable.
%     \end{lstlisting}
%     Charlie and Diveena work in the IT department.
%     When they have checked an app they say it is runnable.
%     \begin{lstlisting}
% "emma" says "charlie" can-act-as "it-department".
% "emma" says "diveena" can-act-as "it-department".
%
% "charlie" says "com.facebook.katana" isRunnable.
% "diveena" says "org.thoughtcrime.securesms" isRunnable.
%     \end{lstlisting}
%
%   \item[Overpriviliged applications]
%     Alice is particularly worried about apps stealing her data.
%     She knows that certain apps need certain permissions.
%     For example a photography app needs access to the camera.
%     She checks each app's permissions carefully for things that seem unusual.
%
%     Alice's policy can be implemented by checking the permissions of the app, using constraints.
%     \begin{lstlisting}
% "alice" says App isRunnable
%    if "permissions-policy" isMetBy(App).
% "alice" says "permissions-policy" isMetBy(App)
%    if App isAnApp
%    where
%       category(App, "Photography"),
%       hasPermission(App, "LOCATION") = false,
%       hasPermission(App, "CAMERA") = true.
%     \end{lstlisting}
%
%     Discovering the \emph{normal} permissions for an app can be done by mining large stores of apps, or apps a user has previously installed.
%     Alternately tools like Stowaway~\cite{Felt:2011kj} could be integrated to check apps.

%   \item[Digital Evidence]
%     In \autoref{ssec:language} we described how digital evidence could be used to decide if policies are met.
%     Suppose a digital evidence generating tool, such as \emph{Evicheck}, can be used to check a policy.
%     For example that no audio can be recorded without the users consent~\cite{Seghir:2014uq}.
%     Using this tool someone generates a proof certificate that this property is met.
%     This might be the developer of the app, a third party certification service, or someone else;
%       when we check the proof we will discover if it is valid or not.
%     \begin{lstlisting}
% "alice" says Anyone can-say 0
%   "recording-consent" shownIsMetBy(Evidence, App).
%
% "skb" says "recording-consent"
%   shownIsMetBy("evicheck://evidence123", App).
%
% "alice" says "recording-consent" isMetBy(App)
%   if "recording-consent" shownIsMetBy(Evidence, App)
%   where
%     evicheckCheckEvidence(Evidence, App) = true.
%     \end{lstlisting}
%     In the example here Alice allows anyone to present her with evidence for her policy.
%     A \ac{skb} offers some evidence, and Alice accepts it as if it checks out.
% \end{description}

>>>>

% Not using NatBib but tired of fixing issues with citations being inserted in that format
\newcommand{\citep}[1]{\cite{#1}}

@@

\begin{abstract}
  It can be difficult to find mobile apps that respect your security and privacy.
  At work, employers rely on employees enforcing company policies about apps at the workplace correctly.
  Users must judge apps by the information shown to them by the store.
  Only 17\% of users pay attention to an apps permissions during installation~\cite{Felt:2012hm} and
  most users do not understand how permissions relate to the capabilities of an app~\cite{Kelley:2012bw}.
  We present AppPAL: a policy language for Android to describe app installation  policies.
  AppPAL goes beyond existing policy enforcement tools, such as Kirin~\cite{Enck:2008vn}, with delegation relationships and the use of static analysis tools.
  We use AppPAL to explore what kinds of policies user's use when installing apps, finding a novel example of the privacy paradox when comparing users privacy views to their actual behaviour.
\end{abstract}

@@

Finding the right apps can be tricky.
Users need to discover which are not going to abuse their data.
This can be difficult as it isn't obvious how apps use the data each has access to.
Consider a user attempting to buy a flashlight app.
By searching the Play store the user is presented with a long list of apps.
Clicking through each one they can find the permissions each requests but not the reasons why each was needed.
They can see review scores from users but not experts who can run tools to check apps for problems and issues like SSL misconfigurations~\cite{Fahl:2012dj}.
If they want to use the app at work will it break their employers rules for mobile usage?

App stores give some information about their apps; descriptions of the app and screenshots as well as review scores.
Android apps show a list of permissions when they're first installed.
Soon new apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use and which permissions to grant must be made by the device user.

Some apps are highly undesirable.
Many \ac{pup} are being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work.
An employer may restrict which apps their employees can use.
The IT department may set a policy---a series of rules describing what kinds of apps may be used and how---to prevent information leaks.
Some users worry apps will misuse their personal data---sending their address book or location to an advertiser without their permission.
Such a user avoids apps which can access their location, or address book.
They may apply their own personal security policies when downloading and running apps.

These policies can only be enforced by the users continuously making the correct decision when prompted about apps.
This is error-prone.
An alternative is to write the policy down and make the computer enforce it.
To implement this we use a logic of authorization---a language designed to express rules about permissible actions.
The policy is written in the logic and enforced by checking the policy is satisfied.

We present AppPAL, an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool} and predicates that allow us to decide which apps to run or install.
The language allows us to reason about apps using statements from third parties.
The implementation allows us to enforce the policies on a device.
We can express trust relationships amongst these parties; use constraints to do additional checks, such as using static analysis results.
This lets us enforce more complex policies than existing tools such as Kirin~\cite{Enck:2009ko} which are limited to permissions checks.

A user, Alice, may have rules she has to follow when using apps for work and her own policies when using apps at home in her private life.
Using AppPAL we can write for work and home, and decide which policy to enforce using a user's location, or the time of day:
\begin{lstlisting}
"alice" says App isRunnable
  if "home-policy" isMetBy(App)
  where at("work") = false.

"alice" says App isRunnable
  if "work-policy" isMetBy(App)
  where beforeHourOfDay("17") = true.
\end{lstlisting}
We can delegate policy specification to third parties or roles, and assign principals to roles:
\begin{lstlisting}
"alice" says "it-department" can-say "work-policy" isMetBy(App).
"alice" says "alice" can-act-as "it-department".
\end{lstlisting}
We can write policies specifying which permissions an app must or must not have by its app store categorization:
    \begin{lstlisting}
"alice" says App isRunnable
   if "permissions-policy" isMetBy(App).
"alice" says "permissions-policy" isMetBy(App)
   if App isAnApp
   where
      category(App, "Photography"),
      hasPermission(App, "LOCATION") = false,
      hasPermission(App, "CAMERA") = true.
    \end{lstlisting}

There has been a great amount of work on developing app analysis tools for Android.
Tools such as Stowaway~\cite{Felt:2011kj} detect over-privileged apps.
TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
ScanDAL~\cite{Kim:2012vt} and SCanDroid~\cite{Fuchs:2009vi} help detect privacy leaks.
Appscopy~\cite{Feng:kPGZr_ja} searches for specific kinds of malware.
Tools like DroidRanger~\cite{Zhou:2012tb} scan app markets for malicious apps.
Many others exist checking and certifying other aspects of app behaviour.

AppPAL can act as a \emph{``glue"} between static analysis tools and the app installation policies device owner's are trying to enforce.
For example a user might not want to install apps with SSL errors (which Mallodroid can check for~\citep{Fahl:2012dj}) or malicious apps as determined by their anti-virus.
Using AppPAL we can combine tools for checking apps to implement the user's policies.
\begin{lstlisting}
"alice" says App isInstallable
  if App isAnApp
  where mallodroidCheck(App) = true,
        avCheck(App) = true.
\end{lstlisting}

\noindent
In this paper we have:
\begin{itemize}
  \item
    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
    and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
  \item
    Described our implementation of the AppPAL language as a library for Android and the JVM.
    We show how the language can describe app properties and Android security policies (\autoref{ssec:language}) as well as how to check policies (\autoref{ssec:eval}).
  \item
    Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
\end{itemize}

@@

An employee \emph{Alice} works for her employer \emph{Emma}.
Emma allows Alice to use her personal phone as a work phone but has some specific concerns.
\begin{itemize}
  \item Alice shouldn't run any apps that can track her movements.
    The testing labs are at a secret location and it mustn't be leaked.
  \item Apps should come from a reputable source, such as the Google Play Store.
  \item Emma uses an \ac{av} program by McAfee.
    It should check all apps before they're installed.
\end{itemize}

To ensure this policy is met Alice promises to follow it.
She might even sign a document promising never to break the rules within the policy.
This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
Alternately Emma's policy could be partially enforced using existing tools.
\emph{Google's Device Policy for Android}~\cite{GoogleAppsDeviceP:tz} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}.

AppPAL builds on these tools, but make the trust and delegation relationships explicit and provide a rigorous and extensible means for deciding whether an app meets a policy.
Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
These could be used to used to ensure no location data is ever obtained.
Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are never leaked to the web.
Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.

@@

% Our research looks at the problem of picking software.
% Given there are some apps you want to install and run and others you do not:
%   how can you express your preferences in such a way that they can be enforced automatically?
% How can we translate policy documents from natural language into a machine checkable form?
% How can we show the trust relationships used to make these decisions clearly and precisely?

@@

We propose a change to the mobile ecosystem, shown in \autoref{fig:ecosystem}.
People have policies which are enforced by AppPAL on their devices.
They can be composed with policies from employers or others to create enhanced devices that ensure apps meet the policies of their owners.
The device can make use of vetting services which run tools to infer complex properties about apps.
Users can buy from enhanced stores which ensure the only apps they sell are the apps which meet the explicitly specified store policies.
Developers could decide which stores to sell their apps in on the basis of policies about stores.

@@

"emma" says "google-play" can-say
  App isReputable. @\label{lst:corporate_4}@
    \end{lstlisting}
  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       firstnumber=15,
                       basicstyle=\scriptsize\ttfamily,
                       stringstyle=\scriptsize\sffamily,
                       keywordstyle=\scriptsize\slshape]
"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = false.

@@

In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isRunnable} (a predicate).
She allows her to delegate the decision further if she chooses (\code{can-say inf}).
Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
If Emma can be convinced all these policies are met then she will say the \code{App isRunnable}.
In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isReputable};
  with \code{"google-play"} specified as the decider of what is buyable or not.
This time Google is not allowed to delegate the decision further (\code{can-say}).
In other words Google is not allowed to specify Amazon as a supplier of apps as well.
Google must say what is buyable directly for Emma to accept it.
Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
Here we use a constraint.
When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
Only if this returns \code{false} will the policy be met.
To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.

@@

AppPAL is implemented as a library for Android and Java.
The parser is implemented using ANTLR4.
The structure of an AppPAL is inherited from SecPAL~\cite{Becker:2006vh} can be seen in \autoref{fig:assertion}.

\begin{figure}
  \newcommand{\bracetext}[1]{\text{\sffamily #1}}
  \newcommand{\smalltext}[1]{\text{\ttfamily\scriptsize #1}}
  \centering
  \begin{minipage}{0.49\linewidth}
    \begin{equation*}\small
      \begin{array}{r l}\footnotesize
        \overbrace{\smalltext{`user'}}^{\bracetext{speaker}} &
        \smalltext{ says }\overbrace{\overbrace{\smalltext{ App }}^{\bracetext{subject}}\overbrace{\smalltext{ isRunnable}}^{\bracetext{predicate}}}^{\bracetext{fact}} \\
        & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
        & \overbrace{\smalltext{ where hasPermission(App, `INTERNET') = true}}^{\bracetext{constraint}}.
      \end{array}
    \end{equation*}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}
  \newcommand{\nonterminal}[1]{$\langle$#1$\rangle$}
  \newcommand{\terminal}[1]{\textbf{#1}}
  \begin{tabular}{r c l}
    \footnotesize
    \nonterminal{Assertion} & $\coloneqq$ & \nonterminal{E} \terminal{says} \nonterminal{Fact} \\
                            &             & \hspace{1em}(\terminal{if} (\nonterminal{Fact}\terminal{,})+)?\terminal{.} \\
                            &             & \hspace{1em}(\terminal{where} \nonterminal{Constraint})? \\
    \nonterminal{Fact}      & $\coloneqq$ & \nonterminal{E} (\terminal{isRunable} $\vert$ $\ldots$) \\
                            & $\vert$     & \nonterminal{E} \terminal{can-say} \terminal{inf}? \nonterminal{Fact} \\
                            & $\vert$     & \nonterminal{E} \terminal{can-act-as} \nonterminal{E} \\
    \nonterminal{E}         & $\coloneqq$ & \terminal{Variable} $\vert$ \terminal{`constant'}
  \end{tabular}
  \end{minipage}
  \caption{Structure and simplified grammar of an AppPAL assertion.}
  \label{fig:assertion}
\end{figure}

In SecPAL the precise nature of predicates and constraints is left open.
In instantiating SecPAL, AppPAL makes the predicates and constraints explicit.
AppPAL policies can make use of the predicates and constraints in \autoref{tab:predicates}.
Additional predicates can be created in the policy files, however constraints are implemented individually.
For example on Android the \texttt{hasPermission} constraint uses the Android package manager to check what permissions an app requests, but the Java version uses the Android platform tools to check.

\begin{table}
  \footnotesize
  \newcommand{\style}[1]{{\footnotesize #1}}
  \begin{tabulary}{\linewidth}{l L}
  \toprule
  \style{Name}                                             & \style{Description}                                                                                                                        \\
  \midrule
  \style{App \texttt{isRunnable}}                          & \style{Says an app can be run.}                                                                                                            \\
  \style{App \texttt{isInstallable}                      } & \style{Says an app can be installed.                                                                                                     } \\
  \style{App \texttt{isAnApp}                            } & \style{Tells AppPAL that an app exists.                                                                                                  } \\
  \style{Policy \texttt{isMetBy(}App\texttt{)}           } & \style{Used to split policies into smaller components.}                                                                                    \\
  \style{\texttt{hasPermission(}App, Permission\texttt{)}} & \style{Constraint to check if an app has a permission.                                                                                   } \\
  \style{\texttt{beforeHourOfDay(}time\texttt{)}         } & \style{Constraint used to check the time.                                                                                                } \\
  \style{Tool\texttt{Check(}App, Property)               } & \style{Constraint to run an analysis tool on an app.                                                                        }              \\
  \bottomrule
\end{tabulary}
\caption{AppPAL predicates and constraints.}
\label{tab:predicates}
\end{table}

@@

In \autoref{sec:problem} and \autoref{sec:idea} we described a \emph{no-tracking-policy} to prevent a user's location being leaked.
In Emma's policy we checked this using the app's permissions.
If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
Some apps may want to access this data, but may not leak it.
We could use a taint analysis tool to detect this (e.g.~FlowDroid~\cite{Fritz:2013vi}).
Our policy now becomes:

@@

"emma" says "location-taint-analysis" isMetBy(App)
  if App isAnApp
  where
    flowDroidCheck(App, "Location", "Internet") = false.
\end{lstlisting}

@@

AppPAL has the same policy checking rules as SecPAL~\cite{Becker:2006vh}.
AppPAL uses an assertion context of known facts and rules
We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based checking algorithm, as no DatalogC library exists for Android.
We have implemented the rules directly in Java.
Pseudo-code is given in \autoref{fig:pseudocode}.

Like Becker~\etal~we make use of an assertion context to store known statements and rules, as well as facts that are deduced during checking.
On a mobile device memory is at a premium.
We would like to keep the assertion context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
This gives us greater control of the evaluation and how the assertion context is created.
For example, when checking the \code{isAnApp} predicate;
  we can fetch the assertion that the subject is an app based on the app in question.

@@
